---
title: "datafest_clean"
author: "Kat Craig,Gloria"
date: "2025-04-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(ggplot2)
```

## Cleaning:

### Data info:

-Savills
- represents tenants (corp, nonprofit, hospitals)
- consult on moving, redesign, etc
- workplace team: advise clients on design
- gather data from surveys, interviews,etc
- space type, size quantities for optimization
- client will understand what they need size wise to move offices

- Consider:
- what is market telling you:
- what are competitors doing
- where are other people living 
-- commute times

- Look at things to differentiate themselves
- looking for trends is important to get ahead of development
- safety very important
- looking for next big spot instead of current ones

What is costar:
- aggregate market data
- important for assessing rent and trends
- Kastle data helps assess attendance in the office to help with optimization
- Can optimize by not including everyone in company in needed office size

Other data:
- census data
- talent demographics for areas
- economic data(stocks, gdp -> office rent, attendance, productivity)

Support clients to help them better use real estate
- oversee local/industry markets

```{r}
leases=read_csv("datafest-data-read-only/Leases.csv")
names(leases)
filter(leases, !is.na(leases$sublet_internal_class_rent))
```

## Including Plots

You can also embed plots, for example:

```{r}
leases_rename <- leases %>%
                rename(sqft=leasedSF, 
                       lease_type=space_type,
                       total_rentable_area_market=RBA,
                       available_rentable_area_market=available_space,
                       available_prop=availability_proportion,     
                       direct_available_rentable_area_market=direct_available_space,
                       direct_available_prop=direct_availability_proportion,
                       sublet_available_area_market=sublet_available_space,
                       sublet_available_prop= sublet_availability_proportion,
                       total_leased_area=leasing
                        ) %>%
  select(-building_id, -building_name,-company_name) %>%
  as.factor(internal_class, )
```
-change quality "A" to factor "High", "O"->"Other"
-transaction type: spread "go" vs "stay"
- move direct_available_space to be 

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
