---
title: "datafest_clean"
author: "Kat Craig, Gloria Yang, Amishi"
date: "2025-04-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(stringr)
library(rvest)
library(curl)
install.packages(curl)
```

## Cleaning:

### Data info:

-Savills
- represents tenants (corp, nonprofit, hospitals)
- consult on moving, redesign, etc
- workplace team: advise clients on design
- gather data from surveys, interviews,etc
- space type, size quantities for optimization
- client will understand what they need size wise to move offices

- Consider:
- what is market telling you:
- what are competitors doing
- where are other people living 
-- commute times

- Look at things to differentiate themselves
- looking for trends is important to get ahead of development
- safety very important
- looking for next big spot instead of current ones

What is costar:
- aggregate market data
- important for assessing rent and trends
- Kastle data helps assess attendance in the office to help with optimization
- Can optimize by not including everyone in company in needed office size

Other data:
- census data
- talent demographics for areas
- economic data(stocks, gdp -> office rent, attendance, productivity)

Support clients to help them better use real estate
- oversee local/industry markets

```{r}
leases=Leases
names(leases)
filter(leases, !is.na(leases$sublet_internal_class_rent))
```

## Including Plots

You can also embed plots, for example:

```{r}
leases1 <- leases %>%
                rename(sqft=leasedSF, 
                       lease_type=space_type,
                       total_rentable_area_market=RBA,
                       available_rentable_area_market=available_space,
                       available_prop=availability_proportion,     
                       direct_available_rentable_area_market=direct_available_space,
                       direct_available_prop=direct_availability_proportion,
                       sublet_available_area_market=sublet_available_space,
                       sublet_available_prop= sublet_availability_proportion,
                       total_leased_area=leasing
                        ) %>%
  select(-building_id, -building_name,-company_name,-address,-costarID)

  leases1$lease_type <- as.factor(leases1$lease_type)
  leases1$internal_class <- as.factor(leases1$internal_class)
  leases1$region <- as.factor(leases1$region)
  leases1$transaction_type <- as.factor(leases1$transaction_type)
  leases1$CBD_suburban <- as.factor(leases1$CBD_suburban)
```
-change quality "A" to factor "High", "O"->"Other"
-transaction type: spread "go" vs "stay"
- move direct_available_space to be 

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r}
write_csv(leases1, "leases_clean.csv")
```

### Joining city data:

```{r}
zips=read_csv("uszips.csv")
zips<-zips %>%
  select(zip, lat,lng) %>%
  rename(long=lng)
zips$zip <- as.double(zips$zip)

leases2 <- leases1 %>%
            left_join(zips, by="zip")
```
```{r}
write_csv(leases2, "leases_coords.csv")
```

```{r}
leases_loc <- leases2 %>%
                      group_by(zip, internal_class) %>%
                      summarise(lat=mean(lat, na.rm=TRUE),
                                long=mean(long, na.rm=TRUE),
                                avg_overall_rent=mean(overall_rent, na.rm=TRUE),
                                avg_tot_leased_area=mean(total_leased_area, na.rm=TRUE),
                                internal_class=internal_class,
                                year=year,
                                avg_prop=mean(available_prop, na.rm = TRUE),
                                avg_sqft=mean(sqft, na.rm=TRUE),
                                count=n()) %>% ungroup()
                        
ggplot(data=filter(leases_loc, avg_sqft<=400000), mapping=aes(x=year, y=avg_sqft,color=internal_class))+ 
  geom_point(alpha=0.2)+
  geom_boxplot()

leases_loc %>%
  ggplot(aes(x=long, y=lat, size=count,color=avg_prop)) +
    borders("state") +
    geom_point() +
    coord_quickmap() +
    scale_color_gradient(low="lightgray", high="red")

leases2%>%filter(zip==10199 |zip == 10041|zip==10121)
leases2 %>% group_by(state, lat, long)%>%
  summarise(avg_sqft=mean(sqft, na.rm=TRUE),
            count=n(),
            lat=mean(lat, na.rm=TRUE),
            long=mean(long, na.rm=TRUE),
            region=region) %>% arrange(desc(avg_sqft)) %>%
  ggplot(aes(x=long, y=lat, shape=region,color=avg_sqft, alpha=0.3)) +
    borders("state") +
    geom_point() +
    coord_quickmap() +
  scale_color_gradient(low="lightgray", high="red")
  
```

```{r}
leases2 <- leases2 %>%
            mutate(transaction_category = recode(transaction_type, New='Go', Relocation='Go', Renewal='Stay',`Renewal and Expansion`='Stay', Restructure='Stay', Extension='Stay', Expansion='Stay', `Sale - Leaseback`='Stay'))
unique(leases2$transaction_category)

leases21<- leases2 %>% group_by(state, transaction_category) %>%
            summarise(count=n(),
                      avg_rent=mean(overall_rent, na.rm=TRUE),
                      prop_cat=count/nrow(filter(leases2, transaction_category==transaction_category)))

            ggplot(data=filter(leases21,transaction_category=='Stay'|transaction_category=='Go'), mapping=aes(x=count, y=state)) +
            geom_col(mapping=aes(fill = transaction_category), position="fill") +
              labs(x="Proportion Stay vs. Go", y="State")+
              guides(fill=guide_legend("Stay or Go"))
```

```{r}
ggplot(data=filter(leases21,transaction_category=='Stay'), mapping=aes(x=avg_rent, y=prop_cat))+
  geom_smooth(mapping=aes(color=transaction_category))+
  coord_cartesian(xlim=c(20,80)) 
ggplot(data=filter(leases21,transaction_category=='Go'), mapping=aes(x=avg_rent, y=prop_cat))+
  geom_smooth(mapping=aes(color=transaction_category)) +
  coord_cartesian(xlim=c(20,80))

ggplot(data=leases21, mapping=aes(x=avg_rent, y=prop_cat))+
  geom_smooth(mapping=aes(color=transaction_category))
```








